Nitrogen cycling in the CLM5: controls on responses to environmental forcing

### 2018-11-13 main.tex

These cases also showed increases in leaf N content for N fertilization, LAI, GPP and NPP. Given that small values of  \emph{cn\_flex\_b} should in fact decrease the amount f carbon spent on N uptake, and thus they

and \emph{cn\_flex\_c}\\
**These will all need looking at once a more careful set of runs are done**

\section{Discussion}
The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).  In all of these activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.  At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.  In particular, as expected, parameters that relate most strongly to the acquisition of N, \emph{frac\_fixers}, \emph{Ncosts} and use \emph{leafcn}, \emph{cn\_flex\_b}, have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.  It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}) . In the CMIP6 generation of land surface models, it is expected that a much larger number of model submissions will contain active representations of N limitation. We do not yet know how CLM5 compares to others in it's carbon cycle feedback strength, but we propose that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained.

\subsection{Limitations}
\subsubsection{Limitations of step-change experiments}
Here investigate the parameter space of the model using step-change experiments. However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions. We assessed the response to the ramping historical CO2, to illustrate how parameter responses differed between the rapid and slow responses. The results are in figure SI **

\subsubsection{Limitations from single site analyses}
The analysis presented here deliberately does not focus on the relative fit of the model to the data collected at the individual sites. The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, which is beyond the scope of this paper. Instead, we chose a set of individual sites, rather than global analysis, to allow for the 936 model integrations that comprise our sensitivity tests, and to probe a set of alternative vegetation and climate conditions. This study is incomplete, as are all investigations of the climatic and parameteric space of land surface models.

\subsubsection{Limitations of parameter calibration issues}
During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the different data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets. Thus, the parameterization used in the default version of the model is a less objective calibration than is ideal, and a limited set of parameters were tuned 'by hand' to avoid excessively spurious values for some model states.  Given the sensitivity of the fertilization responses in the CLM5, we advocate careful interpretation of the outcome of the model, particularly with regard to the climate feedback strength.

\subsubsection{Limitations of the CLM5 revealed by this parameter analysis

1. Lack of optimization of the leaf N area.

\subsection{Future Investigations}
\subsubsection{Temperature and water responses}
Here we probe the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature and moisture changes. The suite of parameters than in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implictly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations. Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by [Dagon et al in prep. and Kennedy et al., in prep.]

\subsubsection{Filtering parameter space with observations}
The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.  In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of availble and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic. Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.

@@

>>>>

Impact of \emph{cn\_flex\_c}\\
The impact of \emph{cn\_flex\_c} (the paremter controlling the degree to which plant CN ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon at Harvard Forest and Niwot Ridge (and Caxiuana, to a smaller extent). These impacts do not mirror those on leaf area index or NPP. At Harvard Forest and Caxiuana, there is a relative increase in the leaf N content, while at Metolius and Niwot, this parameter has a dominant control over the impacts of leaf N to CO$_{2}$ fertilization.  The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high, shown clearly at Niwot Ridge and Metolius (e.g. Figure \ref{NR1 celev}, \ref{MET celev}).  Of course, if N concentrations are allowed to decline too much in response to rising CO$_{2}$, then productivity declines, and other complext feedbacks occur, as shown at Harvard Forest and Niwot Ridge (Figure \ref{HRV celev}, \ref{CAX celev}).  In terms of the overall impact, we can see that the degree of flexibility of the tissue CN ratio acts as an orthoganal control on fertilization response to the total fixation capacity.

\section{Discussion}
The overall responses of land surface models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP model inter-comparison projects and associated activities  (\cite{meehl2014}).  In all of these activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.  At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5 compared to it's predecessors, but it is clear from these investigations that many model parameters can and will have had a large impact on these outcomes.  In particular, as expected, parameters that relate most strongly to the acquisition of N, (\emph{frac\_fixers}, \emph{Ncosts}) and to the flexibility of plant tissues in their C:N ratio (\emph{leafcn}, \emph{cn\_flex\_b}, \emph{cn\_flex\_c}), have dominant roles over the fertilization responses to CO$_{2}$ and nitrogen.  It has long been understood that uncertainties in the representation of Nitrogen are large(\cite{zaehle2014}), but previous model-model comparison studies have struggled to illustrate the exact cause of differences between models predictions, given the complexity of the structural and parametric differences involved. The new CLM5 model has a number of explicit parameterizations, many of which were hard-coded or assumed to be 0 or 1 in previous versions of the model. This allows both a clear exploration of the sensitivities of the parameter space to critical assumptions, as shown here, as well as more comprehensively tuning by future studies and comparisons with relevant datasets.

In the CMIP6 generation of land surface models, it is expected that a greater number of model submissions will contain active representations of N limitation. We propose here that understanding the model parameterizations and structural decisions that contribute to the relative fertilization responses is critical to understanding the implications of the range of results obtained for all models.

\subsection{Limitations}
\subsubsection{Limitations of step-change experiments}
Here investigate the parameter space of the model using step-change experiments. However, such experiments necessarily involve unrealistically rapid changes in environmental boundary condition, and cannot necessarily be considered as analagous to either real-world or transient future scenario conditions. Further, we expect that the carbon and nitrogen cycles mayb take some time to adjust to the new conditions, given slow time-scale processes associated with decomposition of litter with change C:N ratios, and feedbacks between increasing vegetation biomass and soil biogeochemistry.

To test the impact of time on the fertilization responses, we also analyzed the impact of fertilization at 5 and 50 years into the elevated CO$_{2}$ and N experiments. At 50 years, the impacts of fertilization on GPP, NPP and LAI were similar, result not shown) to those at 15 years, but at 5 years, in the N limited sites (NWR and HVF), much more significant impacts of N fertilization were apparent x~2.06 impact on NPP, compared to ~0.95 in the 2015 analysis at NVR), indicating that the results described above had already been subject to progressive N limitation and were closer to an equilibrated state.

\subsubsection{Limitations from single site analyses}
The analysis presented here deliberately does not focus on the relative fit of the model to the data collected at the individual sites. The fit at individual sites of the globally adjusted model cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, which is beyond the scope of this paper.

Instead, we chose a set of individual sites, rather than global analysis to extensively probe a set of alternative vegetation and climate conditions under realistic forcing scenarios.  We also do not make any assessment of how well the model output fits observations across this parameter space, since the relevant datasets with which to test alternative model configurations are, we argue, the set of globally-relevant data products described by \cite{lawrence2018}. Future studies will consider the impact of alternative parameterizations, informed by these site-level analyses, on the global performance of CLM5, but it is clear that a wide range of biogeochemical responses are possible, within the context of this model structure.

\subsubsection{Parameter calibration of the default model, or absence thereof}

During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces,  3) the tendency of the modeled ecosystem to die under past low-CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting pattern that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the different data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets. The parameterization used in the default version of the model utilized the mean trait values where possible, from trait datasets (\cite{lawrence2018}) in combination with targeted tuning of parameters   (including allocation parameters and Nitrogen costs)that were less well constrained by data, to avoid excessively spurious values for some model states.

We recognize that this less objective calibration method is than is ideal, and ongoing studies (\cite{dagon}) will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated in a recent study by \cite{fer2018}, .

The development of the CLM5 slightly pre-dated the operational usage of the International Land Model Benchmarking Project (ILAMB) package \cite{collier2016} for assessing model skill across a very broad range of model components.  In principle, integrating more targeted dataset and data products should narrow the range of model parameter sets that are acceptable in comparison with the suite of available and relevant datasets, as well as accelerate the rejection of model structures that are unrealistic. Ideally, this effort should include synthesis of ecosystem manipulation experiments that provide altered boundary conditions in real life, given the large differences observed here between the sets of parameters affecting present day conditions and those that impact transient responses.

\subsubsection{Temperature and water responses}
In this study we investigate the impacts of parametric uncertainty on the carbon and nitrogen responses of the CLM5. In particular, we probe a targeted parameter set comprising those parameters that determine the impacts of CO$_{2}$ and N fertilization. The full response of the model has considerably more dimensions than these two transient properties. In particular, here we have not considered responses to temperature nor to changes in the supply or demand of water. The suite of parameters than in principle control temperature responses of the model are somewhat distinct from the parameters considered here, given the large number of processes that are directly or implictly affected by temperature (photosynthesis, respiration, soil decomposition, N mineralization, cryosphere interaction) and should be the subject of further investigations. Properties more closely linked to water availability are associated with the hydrological components of the model as well as the PHS plant hydraulics scheme and will be investigated further by \cite{dagon} and \cite{kennedy}.

@@

In this study, we found numerous instances where the model reduced carbon use efficiency by increasing tissue C:N ratios in response to forcing. In principle, there exists an optimal leaf CN ratio that maximizes leaf C export under a given set of conditions. Despite its optimal allocation of leaf N within the leaf (to photosynths
